---
title: "MedfateValidation Docs"
author: "Víctor Granda"
date: "`r Sys.Date()`"
params:
  wd: '/home/malditobarbudo/Documentos/00_Trabajo/CTFC/calibration_validation_medfate/work_structure/Validation'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(root.dir = params$wd)
```

## Load data and build medfate objects

`MedfateValidation` package provides methods to load the sites data and build
the needed `medfate` objects:

### Load data

  - `load_treeData(site)`
  - `load_shrubData(site)`
  - `load_seedData(site)`
  - `load_miscData(site)`
  - `load_meteoData(site)`
  - `load_soilData(site)`
  - `load_soilDataUnilayer(site)`
  - `load_measuredData(site)`
  - `load_terrainData(site)`
  - `load_customParams(site)`
  - `load_remarks(site)`

These functions try to load the corresponding data from the specified site,
indicated by the site code as character (i.e. `'PRADES'`). If the specified
data does not exist (some data are optional) the function will return `NULL`,
as expected by the build functions.

  - `load_rdatas(type = 'Definitive')`

This function allows for load the rdatas containing all the objects produced
along the validation process for all sites. It returns a list of environments
with the objects, that can be accessed as nested lists:  

```{r load_rdatas, eval = FALSE}
objects_reports <- load_rdatas(type = 'Definitive')
objects_reports$`PRADES`$terrainData$elevation
```

`type` argument indicates which reports to read the data from, "Global" for
the global validation, "Temperature" for the temperature validation and
"Definitive" for the five definitive sites.

  > All the load functions are expected to work with the **Validation** folder
    as the root directory. In the case of load_rdatas fuction, it also looks
    for the rdata files generated with the **installed version** of medfate
    (see `packageVersion('medfate')` to know which ones is going to be loaded)

### Build objects

  - `buildForest(treeData, shrubData, seedData, miscData)`
  - `buildSoil(soilData)`

These functions allows to build the `forest` and `soil` objects from `medfate`
package starting on the site data. `treeData`, `miscData` and `soilData` are the
only mandatory arguments, as the remaining can be `NULL` if they are not
available (see sites_data_structure vignette for more info about the site
data objects).

### Examples

If we want to work with the PRADES site, we can load the data and build the
objects to see a summary of them:

```{r load_example}
# load the packages
library(medfate)
library(MedfateValidation)
library(dplyr)

# site/plot code
code <- 'PRADES'

# params
data('newParams')

# raw data
treeData <- load_treeData(code)
shrubData <- load_shrubData(code)
seedData <- load_seedData(code)
customParams <- load_customParams(code)
measuredData <- load_measuredData(code)
meteoData <- load_meteoData(code)
miscData <- load_miscData(code)
soilData <- load_soilData(code)
terrainData <- load_terrainData(code)
remarks <- load_remarks(code)

# model objects
forest_object <- buildForest(treeData, shrubData, seedData, miscData)
soil_object <- buildSoil(soilData)

summary(forest_object, newParams)
soil_object
```

And, in the case of being interested to dive deep in the report for PRADES
site we can load the rdatas and inspect PRADES site:

```{r rdatas_example}
objects_reports <- load_rdatas('Global')
names(objects_reports)

PRADES <- objects_reports$`PRADES`
names(PRADES)
plot(PRADES$res_complex, 'LAI')

# lets remove the objects_reports to avoid conflicts with the following code in
# this vignette
rm(objects_reports)
```

## Parameters

`MedfateValidation` provides a new table of parameters (`newParams`) equivalent
to the `SpParamsMED` provided by `medfate`. In this case, parameters are
obtained from `SpParamsMED` and modificated by those appearing in the HydraTRY
project (courtesy of T. Rosas, M. Mencuccini and JM Vilalta) and the Quercus
monography chapter by EMS Robert, JM Vilalta and M. Mencuccini.
The inclusion here of this table, not open to everyone as the one in `medfate`
is due to the not yet clarified terms of use of TRY derivative works.
For more info about the process of building the `newParams` table, see the
Parametrization vignette.

```{r newParams}
data('newParams', package = 'MedfateValidation')
head(newParams)
```

### Modification of parameters "on the fly"

Along the validation process sometimes is necessary to modify some of the
parameters present in `newParams`, but only for that site in that run. Site data
has a table named `customParams` with the values to modify (see
sites_data_structure vignette). `MedfateValidation` provides a method to modify
`newParams` based on `customParams` called `SpParams_mod`:

```{r params_mod}
# lets modify customParams, as in this site there is no modifications
customParams$xylem_kmax <- c(0.5, 0.8)

# look at the original values of newParams for xylem_kmax
newParams[newParams$Name %in% c('Pinus sylvestris', 'Quercus ilex'), 'xylem_kmax']

# modify the params
sp_params <- SpParams_mod(newParams, customParams)

# check if params have been correctly modified
sp_params[sp_params$Name %in% c('Pinus sylvestris', 'Quercus ilex'), 'xylem_kmax']
```

### Modification of swb.input object "on the fly"

`customParams` can also contain `swb.input` parameters to be modified (i.e. if
the plot has the LAI measured for the cohorts). Instead of using these parameters
calculated by the model, we can modify the `swb.input` by means of the
`inputMod` function. At the moment, only parameters of the sections `above` and
`canopy` from the input can be modified:

```{r inputMod_example}
# lets add some modifications to customParams
customParams$LAI_live <- c(1.1, 2.70)

# create the input object from medfate
control_obj <- medfate::defaultControl()
control_obj$verbose <- FALSE
control_obj$transpirationMode <- 'Simple'
# input object
simple_input <- medfate::forest2swbInput(forest_object, soil_object,
                                         sp_params, control_obj)

# check the input values for above
simple_input$above

# and we can now modify it
simple_input <- inputMod(simple_input, customParams)

# check if the modification has been done correctly
simple_input$above

# lets modify also the "canopy" slot from input object
customParams$gdd <- 250

simple_input$canopy

simple_input <- inputMod(simple_input, customParams)

simple_input$canopy
```


## Saving the results of modelling

`medfate` stores the result of the model run in a `swb` class object which can
be used to plot the model results or to get the different data ouputs
(soil water balance, energy balance...):

```{r res_example, fig.show='hold'}
# run the simple model with the soil and meteo generated previously and a new
# input (as the gdd value is incorrect now)
control_obj <- medfate::defaultControl()
control_obj$verbose <- FALSE
control_obj$transpirationMode <- 'Simple'
# input object
simple_input <- medfate::forest2swbInput(forest_object, soil_object,
                                         sp_params, control_obj)
simple_input <- inputMod(simple_input, customParams)

res <- swb(simple_input, soil_object, meteoData)

# take a look to the class of the res object
class(res)

# lets see the soil water balance for example
head(res$SoilWaterBalance)

# now some plots (SWC, Transpiration)
plot(res, 'Theta')
plot(res, 'ET')
```

To be able to perform the validation of the model, the results must be compared
with the *"real"* data (`measuredData`). To make that easier, `MedfateValidation`
provides a function, `saveRes`, with two main tasks:

  1. Combine the complex and simple main data outputs in one single object, a
     list containing two data.frames, one for each model.
  
  2. Write on disk those data.frames as well as the parameter table used in the
     models, for reproducibility purposes.
     
`saveRes` function allows to save only one model, with the other setted to
`NULL`, which is the expected for the statistics functions that will be explained
later:

```{r saveRes_example}
e_meas <- measuredData %>%
  dplyr::summarize_all(funs(all(is.na(.)))) %>%
  as.logical()

e_measured_coh <- names(measuredData)[!e_meas]

models_dfs <- saveRes(simple_res = res, complex_res = NULL,
                      measured_vars = e_measured_coh, spParams = sp_params,
                      site_code = 'PRADES', write = FALSE)

names(models_dfs)

head(models_dfs$simple)
```

`write = FALSE` argument indicates if the results must be written to disk, useful
when debugging some site and there is no necessity of files to be written.  

### Plant total transpiration in `saveRes`

Despite the model calculates the total transpiration for all cohorts, in the
comparision report it is necessary to compare the measured total transpiration
(the sum of the measured cohorts) with the total modelled **only** for
measured cohorts. In this way, `saveRes` function generates a Eplanttot variable
with only the measured cohorts, using the `measured_vars` argument. It must be a
character vector with the transpiration variables measured.

This will impact the plot and statistics functions, as they report only for
measured cohorts, not model output and results, as they have the total
transpiration for all cohorts modelled.

## Statistics functions

Validation, in a nutshell, consists in comparing the model results with the
field measures in order to detect model logic and parametrization problems.
For that, some statistics need to be generated to be able to compare objectively
the model output.  
`MedfateValidation` calculates three different kinds of statistics, Mean Absolute
Error (MAE), Bias and Correlation (r²). Those are performed with three simple
functions:

  - `MAE_calculator`, which takes two vectors, the measured and the model output
     and calculates the mean difference in absolute value.
  - `bias_calculator`, which takes the measured data and the model output and
     calculates the mean difference, with the sign of the bias.
  - `r_squared_calculator`, which takes the measured data and the model output,
     performs a linear model and retrieve the r² value.
     
```{r stats_bricks}
MAE_calculator(measuredData$SWC, models_dfs$simple$W.1)
bias_calculator(measuredData$SWC, models_dfs$simple$W.1)
r_squared_calculator(measuredData$SWC, models_dfs$simple$W.1)
```

These are the *"basic units"* to build the `statistics_summary` function. This
function performs the three statistics for a given variable (SWC, Total/By cohort
transpiration and Temperature) comparing model output with measured values
and also comparing both models, one to each other, to see differences in
performance.  
To see `statistics_summary` function working at full, lets go to generate first
the results of the complex model:

```{r complex_res, cache=TRUE}
control_obj <- medfate::defaultControl()
control_obj$verbose <- FALSE
control_obj$transpirationMode <- 'Complex'

# input object
soil_object <- buildSoil(soilData)
complex_input <- medfate::forest2swbInput(forest_object, soil_object,
                                          sp_params, control_obj)
  
# input modifications (if any)
complex_input <- inputMod(complex_input, customParams)
  
# model run
res_complex <- medfate::swb(complex_input, soil_object, meteoData,
                            terrainData$latitude, terrainData$elevation,
                            terrainData$slope, terrainData$aspect)

models_dfs <- saveRes(res, res_complex, e_measured_coh,
                      sp_params, 'PRADES', write = FALSE)
```

Now `models_dfs` has results of both models stored and `statistics_summary` can
work with both:

```{r stats_summary, results='asis'}
SWC_stats <- statistics_summary(var = 'SWC', models = models_dfs,
                                measured_data = measuredData, soil = soil_object,
                                meteo_data = NULL, trunc = NULL)

knitr::kable(SWC_stats)
```

In this case the statistics for soil water content are showed. The `var` argument
tells the function which stastistics must be performed. The values allowed are:

  - 'SWC': This var needs the soil object to work properly
  - 'Eplanttot': This var only needs the models and the measured data and instead
     of a table, generates a list.
  - 'E_by_Cohort': This var only needs the models and the measured data and instead
     of a table, generates a list.
  - 'Temperature': This var needs the meteo data to work properly, and can use the
    `trunc` argument to subset the data to compare.

```{r stats_summary_2}
Eplanttot_stats <- statistics_summary('Eplanttot', models_dfs, measuredData,
                                      soil_object)

str(Eplanttot_stats)
```

```{r stats_summary_3, results='asis'}
library(dplyr)

Ecohorts_stats <- statistics_summary('E_by_Cohort', models_dfs, measuredData,
                                     soil_object)

rbind(
  round(Ecohorts_stats[["Esp_MAE_simple"]], 5),
  round(Ecohorts_stats[["Esp_r_sq_simple"]], 5),
  round(Ecohorts_stats[["Esp_bias_simple"]], 5),
  round(Ecohorts_stats[["Esp_MAE_complex"]], 5),
  round(Ecohorts_stats[["Esp_r_sq_complex"]], 5),
  round(Ecohorts_stats[["Esp_bias_complex"]], 5),
  round(Ecohorts_stats[["Esp_MAE_both"]], 5),
  round(Ecohorts_stats[["Esp_r_sq_both"]], 5),
  round(Ecohorts_stats[["Esp_bias_both"]], 5)
) %>%
  cbind(
    c('MAE Simple vs. Measured',
      'rsq Simple vs. Measured',
      'Bias Simple vs. Measured',
      'MAE Complex vs. Measured',  
      'rsq Complex vs. Measured',  
      'Bias Complex vs. Measured',
      'MAE Simple vs. Complex',
      'rsq Simple vs. Complex',
      'Bias Simple vs. Complex'),
    .
  ) %>%
  as.data.frame() %>%
  knitr::kable(col.names = c('Statistics', Ecohorts_stats[['Cohort_name']]))

```

  > Temperature stats must be done in other set of sites (Albert Vila plots)
    as they are the only ones with measured canopy temperature


## Visualization

`medfate` provides custom `plot` methods for visualize the results of the model,
but for the validation process we need new functions that generates the plots
comparing the measured data and the modelled data. `MedfateValidation` provides
several functions that are in charge of this visualization task, for each
combination of model type and variable:

  - `plot_swc_simple_gg`
  - `plot_swc_complex_gg`
  - `plot_swc_both_gg`
  - `plot_eplanttot_simple_gg`
  - `plot_eplanttot_complex_gg`
  - `plot_eplanttot_both_gg`
  - `plot_cohort_simple_gg`
  - `plot_cohort_complex_gg`
  - `plot_cohort_both_gg`

These functions are intended to being used inside of the wrapper function
`plot_res_gg`, which is feeded with the variable and models to plot and
generate the visualization:

```{r swc_plot, fig.width=8, fig.height=10.5, warnings = FALSE}
plot_res_gg(variable = 'SWC', models = models_dfs, soil = soil_object,
            measured_data = measuredData, mode = 'both')
```

`variable` argument values are the same as in `statistics_summary` function (except
for the 'Temperature' as it has its own method), and the `mode` argument expects
one of the following:

  - 'both': Plot the simple, complex and the comparision of both models
  - 'simple': Plot the simple model
  - 'complex': Plot the complex model

```{r plots_etot, fig.width=8, fig.height=10.5}
plot_res_gg(variable = 'Eplanttot', models = models_dfs, soil = soil_object,
            measured_data = measuredData, mode = 'both')
```

  > When plotting cohort transpiration, the function returns a list of plots,
    therefore the use of purrr::walk to plotting all in one

```{r plots_ecoh, fig.width=8, fig.height=3.5}
plots_cohorts <- plot_res_gg('E_by_Cohort', models_dfs, soil_object,
                             measuredData, 'both')

purrr::walk(
  plots_cohorts,
  ~ print(.x)
)
```

### SWC by layers

`medfate` can plot the SWC by layer with its custom `plot` method, but
`MedfateValidation` also have a method to perform this, generating both models
plot side by side to easy comparision with the function `plot_swc_layers_gg`:

```{r swc_layers, fig.width=8}
plot_swc_layers_gg(models_dfs)
```

## Reports

Validation process is done by means of html reports generated with Rmarkdown.
This allows for running the code only once, generating the output needed to
visualize the validation site by site, but also allowing for debugging sites
simply using the code contained in the Rmd templates and the RData file
generated with all the objects created in the report. These templates are located
in the folder where the package is installed
(`MedfateValidation/inst/Rmd_templates`). To generate the reports, `MedfateValidation`
provides some wrappers and a internal function. The wrappers are:

  - `global_process`
  - `global_kmax_factor_process`
  - `transpiration_process`
  - `temperature_process`

Each of this wrappers generate a different kind of report. Global processes
generates the global validation process (SWC and Transpiration), with or without
the kmax fixed factor to reduce the transpiration. Transpiration process generates
the report for transpiration taking the real SWC from the measuredData as well as
using a simpler soil conception (only one layer). And, finally, temperature process
runs the report to validate the canopy temperatures.  
All these wrappers share the same arguments:

  - `sites` a vector with site codes to run
  - `wd` the path to the `Validation` folder where all the validation is done
  - `transpMode` a character indicating the models to run, simple, complex or both
  - `SPParams` a character indicating if it will use the "old" SpParamsMED table or
     the  "new" newParams table.
  - `tapering` a logical indicating if the `control` parameter `taper` must be
     performed (default) or not.
  - `rhizosphere` a numeric input indicating the average rhizosphere resistance
     in proportion. This value is used to change the control object from
     default value to the desired value.

All the functions call internally the `report_render` function, a customization of
`rmarkdown::render`.

