---
title: "Transpiration report for `r params$code`"
author: "medfateTeam"
date: "`r Sys.time()`"
params:
  wd: ''
  code: ''
  transpMode: 'both'
  SPParams: 'old'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = params$wd)
library(medfate)
library(MedfateValidation)
library(dplyr)
```

```{r data, message=FALSE, warning=TRUE}
# site/plot code
code <- params$code

# raw data
treeData <- load_treeData(code)
shrubData <- load_shrubData(code)
seedData <- load_seedData(code)
customParams <- load_customParams(code)
measuredData <- load_measuredData(code)
meteoData <- load_meteoData(code)
miscData <- load_miscData(code)
soilData <- load_soilDataUnilayer(code) # soil data for only one layer
terrainData <- load_terrainData(code)

# model objects
forest_object <- buildForest(treeData, shrubData, seedData, miscData)

# SpParams
# check if new or old will be used
if (params$SPParams == 'old') {
  # load
  data('SpParamsMED', package = 'medfate')
  # modify
  sp_params <- SpParams_mod(SpParamsMED, customParams)
}

if (params$SPParams == 'new') {
  # load
  data('newParams', package = 'MedfateValidation')
  # modify
  sp_params <- SpParams_mod(newParams, customParams)
}

# Warning for custom FC
if (!all(is.na((soilData[['FC']])))) {
  warning('Custom FC values has been provided: ', soilData[['FC']][1], ' ...')
}
```

# General Info

```{r gen_info, results='asis'}
knitr::kable(
  data.frame(
    Pkg_version  = as.character(packageVersion('medfate')),
    Site = params$code,
    Date = format(Sys.Date())
  )
)
```

```{r simple_model, eval=TRUE, message=FALSE}
# check if simple mode has to be performed
if (params$transpMode %in% c('simple', 'both')) {
  # control object
  control_obj <- medfate::defaultControl()
  control_obj$verbose <- FALSE
  control_obj$transpirationMode <- 'Simple'
  
  # input object
  simple_input <- medfate::forest2swbInput(forest_object, buildSoil(soilData),
                                           sp_params, control_obj)
  
  # input modifications (if any)
  simple_input <- inputMod(simple_input, customParams)
  
  # model run day by day
  res_simple <- transp_day_by_day(simple_input, meteoData, soilData,
                                  terrainData, measuredData)
} else {
  # if simple mode is not performed, report about it
  res_simple <- NULL
  cat('Simple model has not been selected to be performed')
}
```

```{r complex_model}
# check if complex mode has to be performed
if (params$transpMode %in% c('complex', 'both')) {
  # control object
  control_obj <- medfate::defaultControl()
  control_obj$verbose <- FALSE
  control_obj$transpirationMode <- 'Complex'
  
  # input object
  # we also rebuild the soil_object to avoid using W data from the previous
  # runs
  complex_input <- medfate::forest2swbInput(forest_object, buildSoil(soilData),
                                            sp_params, control_obj)
  
  # input modifications (if any)
  complex_input <- inputMod(complex_input, customParams)
  
  # model run day by day
  res_complex <- transp_day_by_day(complex_input, meteoData, soilData,
                                   terrainData, measuredData)
} else {
  # if simple mode is not performed, report about it
  res_complex <- NULL
  cat('Complex model has not been selected to be performed')
}
```

```{r joining_res}
models_dfs <- list(
  simple = res_simple,
  complex = res_complex
)
```

## Models comparision

### Total Tranpiration

```{r statistics_Eplanttot}
Eplanttot_stats <- statistics_summary('Eplanttot', models_dfs, measuredData,
                                      soil = buildSoil(soilData))
```

Comparision | MAE | r² | Bias
------------|-----|----|-------
Simple *vs.* Measured | `r Eplanttot_stats$Eplanttot_MAE_simple` | `r Eplanttot_stats$Eplanttot_r_sq_simple` | `r Eplanttot_stats$Eplanttot_bias_simple`
Complex *vs.* Measured | `r Eplanttot_stats$Eplanttot_MAE_complex` | `r Eplanttot_stats$Eplanttot_r_sq_complex` | `r Eplanttot_stats$Eplanttot_bias_complex`
Simple *vs.* Complex | `r Eplanttot_stats$Eplanttot_MAE_both` | `r Eplanttot_stats$Eplanttot_r_sq_both` | `r Eplanttot_stats$Eplanttot_bias_both`


```{r plots_Eplanttot, fig.height=6, fig.width=8}
print(plot_res_gg('Eplanttot', models_dfs, buildSoil(soilData),
                  measuredData, params$transpMode))
```

### Transpiration by cohorts

```{r statistics_E_cohorts, results='asis'}
Ecohorts_stats <- statistics_summary('E_by_Cohort', models_dfs, measuredData,
                                     buildSoil(soilData))

rbind(
  round(Ecohorts_stats[["Esp_MAE_simple"]], 5),
  round(Ecohorts_stats[["Esp_r_sq_simple"]], 5),
  round(Ecohorts_stats[["Esp_bias_simple"]], 5),
  round(Ecohorts_stats[["Esp_MAE_complex"]], 5),
  round(Ecohorts_stats[["Esp_r_sq_complex"]], 5),
  round(Ecohorts_stats[["Esp_bias_complex"]], 5),
  round(Ecohorts_stats[["Esp_MAE_both"]], 5),
  round(Ecohorts_stats[["Esp_r_sq_both"]], 5),
  round(Ecohorts_stats[["Esp_bias_both"]], 5)
) %>%
  cbind(
    c('MAE Simple vs. Measured',
      'rsq Simple vs. Measured',
      'Bias Simple vs. Measured',
      'MAE Complex vs. Measured',  
      'rsq Complex vs. Measured',  
      'Bias Complex vs. Measured',
      'MAE Simple vs. Complex',
      'rsq Simple vs. Complex',
      'Bias Simple vs. Complex'),
    .
  ) %>%
  as.data.frame() %>%
  knitr::kable(col.names = c('Statistics', Ecohorts_stats[['Cohort_name']]))
```

```{r plots_Eplant_coh, fig.height=3, fig.width=7}
plots_cohorts <- plot_res_gg('E_by_Cohort', models_dfs, buildSoil(soilData),
                             measuredData, params$transpMode)

purrr::walk(
  plots_cohorts,
  ~ print(.x)
)
```

## Soil Info

```{r soil_info}
if (params$transpMode %in% c('simple', 'both')) {
  print(buildSoil(soilData))
  simple_input[['below']]
} else {
  print(buildSoil(soilData))
  complex_input[['below']]
}
```

## SoilWaterBalance Input

```{r swb_input}
if (params$transpMode == 'simple') {
  print(simple_input)
} else {
  if (params$transpMode == 'complex') {
    print(complex_input)
  } else {
    print(simple_input)
    print(complex_input)
  }
}
```


## Session Info

```{r session_info}
# saving objects
file_name <- file.path('Output', packageVersion('medfate')[[1]],
                       params$code,
                       paste0(format(Sys.time(), "%Y%m%d_%H%M"),
                              '_', params$code, '_',
                              'transp_report_objects.RData'))
save(list = ls(all.names = TRUE), file = file_name)
# printing session info
devtools::session_info()
```

